\section*{Insert Scores Inline}

With the \option{insert=inline} option it is simple to insert arbitrary
notational fragments in the \lilypond{ e'8 d'16 e' } continuous text of a
document. By default the staffsize is scaled to be 2/3 of the staffsize a
regular score would have at this point. This means if the \option{staffsize}
option is modified globally or locally then the staffsize of the inline score is
changed too.

In order to make the size of inline scores independent from the regular
staffsize the option \option{inline-staffsize} can be used the same way as
\option{staffsize}. \lilypond[inline-staffsize=8]{ e'8 d'16 e' } has the inline
staffsize manually set to \texttt{8}.

\paragraph{Alignment and padding} By default inline scores are vertically
centered to a line 1/2em above the text's baseline. \lilypond[valign=top]{ e'8
d'16 e' } but the score can also be aligned \lilypond[valign=bottom]{ e'8 d'16
e' } to the top or the baseline of the text.

Unfortunately this can only consider the borders of the \emph{image} and not
those of the \emph{score} or the staff lines. To alleviate this situation a
specific vertical offset can be given with \option{voffset=-3pt} (or any other
\TeX\ lengths). This offset is calculated after the alignment.
\lilypond[valign=bottom,voffset=-6pt]{ e'8 d'16 e' } is inserted with
\option{valign=bottom,voffset=-4pt}.

Horizontally inline scores are padded by \option{hpadding=0.75ex} -- except if
they happen to appear at the beginning or end of a line, as can be seen in the
last score in the previous paragraph. \lilypond[hpadding=2em]{ e'8 d'16 e' }
Increasing the \option{hpadding} will ensure more space around the score.

\paragraph{Bare Inline scores} \option{insert=bare-inline} will remove all the
staff elements (staff symbol, time signature, clef) by implicitly applying
\option{nostaff}, which is most useful for including notational symbols like
characters in the paragraph.
\lilypond[insert=bare-inline,%
inline-staffsize=14,%
voffset=-1pt,%
hpadding=0.25ex%
]{s2.\startTrillSpan s4\stopTrillSpan}
This actually works like the \option{lilyglyphs}
package\footnote{\url{https://github.com/uliska/lilyglyphs}} but with the
possibility of inserting arbitrary LilyPond material without having to prepare
precompiled PDF images.
