\documentclass{lyluatexexample}

\begin{document}

\VerbatimFootnotes

\section*{Wrapping Commands}

\subsection*{Command within commands}

\cmd{lily} can be wrapped within another command as usual:

\begin{verbatim}
\newcommand\mylily[2][1]{\lily[inline-staffsize=10, #1]{#2}}

This is \mylily[voffset=10pt]{a' b' c''} an example.
\end{verbatim}

\newcommand\mylily[2][1]{\lily[inline-staffsize=10, #1]{#2}}

This is \mylily[voffset=10pt]{a' b' c''} an example.\par

\subsection*{Environment within environments}

\emph{It isn't possible to wrap \highlight{ly} environment within a command.}\par

It's possible to wrap \highlight{ly} within and environment, but there are
several drawbacks\footnote{%
Those drawbacks are:
\begin{itemize}
  \item this custom environment cannot have optional parameters. To be more
  precise, if it has only optional parameters, it will be necessary to add \verb`[]`
  after \verb`\begin{MY_ENV}` if no parameter is specified ; so they're not
  optional any more…
  \item to call \highlight{ly}, you'll have to:
  \begin{itemize}
    \item either write \verb`\begin{ly}[] \end{ly}` (which works with
    \verb`\begin{lilypond}[]` \verb`\end{lilypond}` too) ;
    \item or use the \TeX\ primitives \verb`\ly \endly` (not only for \highlight{ly},
    but also for other environments).
  \end{itemize}
\end{itemize}%
}.

To avoid those drawbacks, \lyluatex\ defines a special command, \verb`\lynewenvironment`,
that behaves as you'd expect from \verb`\newenvironment`.

\begin{verbatim}
\lynewenvironment{myly}{%
  This is \emph{my} lilypond environment.
  \begin{ly}%
}{%
  \end{ly}
}

\begin{myly}
  a b c
\end{myly}
\end{verbatim}

\newenvironment{myly}{%
  This is \emph{my} lilypond environment.
  \begin{ly}%
}{%
  \end{ly}
}

\begin{myly}
  a b c
\end{myly}

\begin{verbatim}
\lynewenvironment{lyfigure}[2][]{%
\edef\mycaption{#2}
\begin{figure}
\begin{center}
  \begin{lilypond}[#1]%
}{%
  \end{lilypond}
  \caption{\mycaption}
\end{center}
\end{figure}
}

\begin{lyfigure}{caption}
a' b' c
d' e' f
\end{lyfigure}
\end{verbatim}

\lynewenvironment{lyfigure}[2][]{%
\edef\mycaption{#2}
\begin{figure}
\begin{center}
  \begin{lilypond}[#1]%
}{%
  \end{lilypond}
  \caption{\mycaption}
\end{center}
\end{figure}
}

\begin{lyfigure}{caption}
a' b' c
d' e' f
\end{lyfigure}

\begin{verbatim}
\lynewenvironment{lyotherfigure}[1][]{%
\edef\option{#1}
\figure
\center
  \ly
}{%
  \endly%
  \def\empty{}\ifx\option\empty\else\caption{\option}\fi
\endcenter
\endfigure
}

\begin{lyotherfigure}
d' e' f
a' b' c
\end{lyotherfigure}
\end{verbatim}

\lynewenvironment{lyotherfigure}[1][]{%
\edef\option{#1}
\figure
\center
  \ly
}{%
  \endly%
  \def\empty{}\ifx\option\empty\else\caption{\option}\fi
\endcenter
\endfigure
}

\begin{lyotherfigure}
d' e' f
a' b' c
\end{lyotherfigure}

\textbf{Important note:} \verb`\lynewenvironment` is intended to insert \emph{LaTeX} code before
and after the scores; due to the special behavior of \verb`ly` environment, it isn't possible
to insert \emph{LilyPond} code that way. So this won't work:

\begin{verbatim}
  \lynewenvironment{myly}{%
    \begin{ly}
      a b c
  }{%
    \end{ly}
  }
\end{verbatim}

To do such a thing, \lyluatex\ defines a command and four options:
\begin{itemize}
  \item \verb`\lysavefrag` lets one save a LilyPond fragment to be re-used afterward;
  \item \verb`include_header`, \verb`include_footer`, \verb`include_before_body` and \verb`include_after_body` options
        let one insert such fragments at designed places within inserted score.
\end{itemize}

So this works:

\begin{verbatim}
\begin{lysavefrag}{head}
a b c
\end{lysavefrag}

\begin{lysavefrag}{foot}
g a' b
\end{lysavefrag}

begin{ly}[
  include_before_body={head,head},
  include_after_body=foot,
]
d e f
\end{ly}
\end{verbatim}

It's also possible to use \verb`\lynewenvironment` to wrap such a command:

\begin{verbatim}
\begin{lysavefrag}{head}
a b c
\end{lysavefrag}

\begin{lysavefrag}{foot}
g a' b
\end{lysavefrag}
  
\lynewenvironment{yourly}[1][]{%
  {\centering test \par}
  \begin{ly}[
    include_before_body={head,head},
    include_after_body=foot,
  ]
}{
  \end{ly}
}

\begin{yourly}
d e f
\end{yourly}
\end{verbatim}

\begin{lysavefrag}{head}
a b c
\end{lysavefrag}

\begin{lysavefrag}{foot}
g a' b
\end{lysavefrag}
  
\lynewenvironment{yourly}[1][]{%
  {\centering test \par}
  \begin{ly}[
    include_before_body={head,head},
    include_after_body=foot,
  ]
}{
  \end{ly}
}

\begin{yourly}
d e f
\end{yourly}

\end{document}
